---
title: ''
author: ''
date: ''
output: 
  html_document:
    df_print: paged
    keep_md: true
  word_document:
    reference_docx: style.1.docx
---

```{r setup, include = FALSE}

  knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                        fig.width  = 6 * 1.67, fig.height = 6)

  library(tidyverse)
  library(viridis)
  library(TOSTER)
  library(rms)
  library(pander)
  library(broom)
  library(geepack)
  library(ggbeeswarm)
  library(flextable)

  load("data/data.RData")
  source("scripts/functions.R")
  
```


# Statistical methods

Categorical patient characteristics were were described by their counts and percentages in each category. Continuous characteristics were described by their means and SDs, medians and IQRs, and total ranges. 

VTE rates and 95% confidence intervals (CIs) were estimated using Poisson regression.

We evaluated the equivalence (and thus non-inferiority) of *`r levels(data$vte_prophylaxis)[25]`* to *`r levels(data$vte_prophylaxis)[24]`* based on the risk difference for VTE, using a margin of 1.0%. We reported both the unadjusted and adjusted risk differences and respective 90% confidence intervals (which is the equivalent of two one-sided tests [TOST] with alpha = 0.05 (REF)). Covariate adjustment was made using the standardized risk difference estimated using the marginal structural binomial regression model outlined in (REF). The covariates were age (years), sex, body mass index (BMI; kg/m2), any history of VTE (none vs DVT vs PE), and procedure (TKR vs THR). Patients with missing covariate data were excluded from the adjusted risk difference models. 

In addition to the head to head comparison between *`r levels(data$vte_prophylaxis)[25]`* and *`r levels(data$vte_prophylaxis)[24]`*, we similarly evaluated the equivalence in the VTE risk difference between patients seen from *2010 to 2012* and those seen from *2013 to 2016*, irrespective of the actual VTE prophylaxis received. This was because *`r levels(data$vte_prophylaxis)[25]`* was almost always used from 2013 on, and very rarely used before 2013 (see supplemental figure 1).   

All analysis were conducted using the R Project for Statistical Computing v3.4.3 (REF). 

# Results

```{r table_1}

# Key variables
  table_1 <- select(
    data, sex, age, over_65, bmi, obese, stay_length, procedure, vte_type, 
    anticoagulant_type_2, dvt_hx, vte_prophylaxis_2
    )

  data_frame(
    Variable       = name.1(table_1),
    N              = n.miss(table_1),
    "Mean SD"      = summary.1(table_1),
    "Median [IQR]" = med.iqr(table_1),
    "(Min, Max)"   = min.max(table_1)
    ) %>% flextable()

  
```

The sample included `r nrow(data)` patients seen between `r min(data$procedure_year)` and `r max(data$procedure_year)`. The median age was `r round(mean(data$age), 1)` years (range = `r min(data$age)` to `r max(data$age)`). 

```{r yearly_incidence}

# Percentage with VTE, over all years
  overall <- round(table(data$vte_outcome)["Yes"] /
                     sum(table(data$vte_outcome)), 5) * 100

# Percentage with VTE in each year, with 95% CIs
  counts <- group_by(data, procedure_year) %>%
    summarise(pop = n(),
              vte_count = sum(as.numeric(vte_outcome) - 1),
              vte_pct = round(mean(as.numeric(vte_outcome) - 1), 4) * 100,
              incidence = signif(vte_count/pop*100, 4),
              se = 1/sqrt(vte_count),
              ll = signif(exp(log(incidence) - 1.96 * se), 4),
              ul = signif(exp(log(incidence) + 1.96 * se), 4),
              ci = paste(ll, " to ", ul)) %>%
    mutate(mean_incidence = overall)

```


The overall percentage of patients experiencing a VTE (2010 to 2016) was `r round(table(data$vte_outcome)["Yes"] / sum(table(data$vte_outcome)), 4) * 100`%. The highest rate of `r max(counts$vte_pct)`% was in `r counts$procedure_year[counts$vte_pct == max(counts$vte_pct)]`, while the lowest rate of `r min(counts$vte_pct)`% was in `r counts$procedure_year[counts$vte_pct == min(counts$vte_pct)]`. VTE rates and 95% CIs across study all years are given in Figure 1. 

*Figure 1. Yearly VTE incidence (95% CIs), regardless of VTE prophylaxis.*

```{r vte_by_year_plot}

   ggplot(counts, 
          aes(y = incidence, x = procedure_year, ymax = ul, ymin = ll)) +
     geom_hline(yintercept = max(counts$incidence), size = 1,
                alpha = 0.3, linetype = "dashed") +
     geom_hline(yintercept = min(counts$incidence), size = 1,
                alpha = 0.3, linetype = "dashed") +
     geom_hline(aes(yintercept = overall), linetype = "dashed", color = "red") +
     geom_pointrange(size = 1) +
     geom_line(size = 1) +
     theme_minimal() +
     scale_x_continuous(breaks = 2010:2016) +
     scale_color_viridis(guide = FALSE, discrete = TRUE) +
     xlab("Year") +
     ylab("VTE cases per 100 patients") +
     ylim(0, 3)
     
  ggsave("plots/f1_yearly_incidence.pdf", device = "pdf", dpi = 600)


```

```{r tost_by_treatment}

# Non-equivalance with the two one-sided test procedure

# Data set to let us model effects of different treatments
  df_tx <- data %>%
    filter(vte_prophylaxis %in% levels(vte_prophylaxis)[24:25] &
           !is.na(vte_prophylaxis)) %>%
    droplevels() %>%
    mutate(vte_prophylaxis_num = as.numeric(vte_prophylaxis) - 2) %>%
    select(vte_outcome_num, vte_outcome, vte_prophylaxis, 
           vte_prophylaxis_num, sex, bmi, age, id, procedure, dvt_hx) 

# Summary data needed to use the TOST function  
  s_1 <- group_by(df_tx, vte_prophylaxis) %>%
    summarise(n = n(), prop = mean(as.numeric(vte_outcome) - 1, na.rm = TRUE))

# TOST  
  tost_1 <- TOSTtwo.prop(
    s_1$prop[2], s_1$prop[1], s_1$n[2], s_1$n[1], -0.01, 0.01, 0.05, 
    plot = FALSE
    )
  
```

```{r adjusted_by_treatment}

# These are the covariate adjusted standardized risk difference estimated with
# marginal structural binomial regression models. REF:
# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6372130/

  df_tx <- df_tx[complete.cases(df_tx), ] 

# Create the IPWs      
  exposed <- glm(vte_prophylaxis ~ age + sex + bmi + dvt_hx + procedure, 
                 family = binomial(link = "logit"), data = df_tx)
  pred_exp <- predict(exposed, type = "response")
  
# Stabilize the weights by multiplying each weight by the overall unadjusted
# probability of that person having exposure level E = e
  sptw <- 
    df_tx$vte_prophylaxis_num * mean(df_tx$vte_prophylaxis_num) / pred_exp + 
    (1 - df_tx$vte_prophylaxis_num) * (1 - mean(df_tx$vte_prophylaxis_num)) / 
    (1 - pred_exp)
  
# Fit linear binomal model to weighted data for E-D association with robuse SEs
  rd_model <- geeglm(
    vte_outcome_num ~ vte_prophylaxis_num, 
    family = binomial(link = "identity"), 
    weight = sptw, id = id, data = df_tx
    ) %>% tidy()
   
  rd <- signif(rd_model[2, 2] * 100, 2)
  ll <- signif((rd_model[2, 2] - (1.64 * rd_model[2, 3])) * 100, 2)
  ul <- signif((rd_model[2, 2] + (1.64 * rd_model[2, 3])) * 100, 2)

```

The unadjusted VTE risk difference comparing *`r levels(data$vte_prophylaxis)[25]`* to *`r levels(data$vte_prophylaxis)[24]`* was `r signif(tost_1$dif * 100, 2)`% with a 90% TOST CI of `r signif(tost_1$LL_CI_TOST * 100, 2)`% to `r signif(tost_1$UL_CI_TOST * 100, 2)`%, which suggests equivalence between the two. The adjusted results were not appreciably different (RD `r rd`% (90% TOST CI `r ll` to `r ul`%)).


```{r tost_by_year}

  df_year <- select(
    data, vte_outcome_num, vte_outcome, year_group, year_group_num, sex, bmi, 
    age, id, procedure, dvt_hx
    )

  s_2 <- group_by(df_year, year_group) %>%
    summarise(n = n(), prop = mean(as.numeric(vte_outcome) - 1, na.rm = TRUE))
  
  tost_2 <- TOSTtwo.prop(
    s_2$prop[2], s_2$prop[1], s_2$n[2], s_2$n[1], -0.01, 0.01, 0.05, 
    plot = FALSE
    )
  
```


```{r adjusted_by_year}

  df_year <- df_year[complete.cases(df_year), ] 

# Create the IPWs      
  exposed <- glm(year_group ~ age + sex + bmi + dvt_hx + procedure, 
                 family = binomial(link = "logit"), data = df_year)
  pred_exp <- predict(exposed, type = "response")
  
# Stabilize  
  sptw <- 
    df_year$year_group_num * mean(df_year$year_group_num) / pred_exp + 
    (1 - df_year$year_group_num) * (1-mean(df_year$year_group_num)) / 
    (1 - pred_exp)
  
# Fit linear binomal model to weighted data for E-D association with robuse SEs
  rd_model_2 <- geeglm(
    vte_outcome_num ~ year_group_num, 
    family = binomial(link = "identity"), 
    weight = sptw, id = id, data = df_year
    ) %>% tidy()
   
  rd_2 <- signif(rd_model_2[2, 2] * 100, 2)
  ll_2 <- signif((rd_model_2[2, 2] - (1.64 * rd_model_2[2, 3])) * 100, 2)
  ul_2 <- signif((rd_model_2[2, 2] + (1.64 * rd_model_2[2, 3])) * 100, 2)

``` 

The unadjusted VTE risk difference comparing patients seen in *2010-2012* to those seen in *2013-2016* was `r signif(tost_2$dif * 100, 2)`% with a 90% TOST CI of `r signif(tost_2$LL_CI_TOST * 100, 2)`% to `r signif(tost_2$UL_CI_TOST * 100, 2)`%, which suggests equivalence between the two. The adjusted results were not appreciably different (RD `r rd_2`% (90% TOST CI `r ll_2` to `r ul_2`%)).


# Demographic predictors

```{r demographic_predictors}

  demo <- glm(vte_outcome_num ~ age + sex + bmi +
                any_dvt_hx, 
              family = binomial(link = "logit"), data = data)

  demo2 <- glm(vte_outcome_num ~ age + sex + bmi +
                dvt_hx, 
              family = binomial(link = "logit"), data = data)

  sjPlot::tab_model(demo, file = "tables/demo.html")
  
  sjPlot::tab_model(demo2, file = "tables/demo2.html")

```


# Supplemental material

*Supplemental Figure 1. Among the most commonly used types of VTE prophylaxis, which were used each year? (2010 to 2016)*
```{r prophylaxis_types_year_plot}

  filter(data, vte_prophylaxis %in% levels(vte_prophylaxis)[20:25] &
         !is.na(vte_prophylaxis)) %>%
  ggplot(aes(x = factor(procedure_year), fill = vte_prophylaxis)) +
    geom_bar(position = "fill") +
    theme_minimal() +
    scale_fill_viridis("", discrete = TRUE, direction = -1) +
    theme(panel.grid = element_blank()) + 
    xlab("") +
    ylab("Proportion of patients") +
    ggtitle("")

  ggsave("plots/sf1_vte_types_year.pdf", device = "pdf", dpi = 600)

```

```{r code_book}

  print(summarytools::dfSummary(data), method = "render")

```

